---
title: "Regression Analysis on Concrete Compressive Strength Data Set"
author: "1. Balaji Baskaran, 2. Nansita Subramanian, 3. Raunak Avinash Deshmukh"
date: "December 6, 2015"
output: html_document
---

```{r, echo=FALSE}
#let the required libraries go in here
library(corrplot)

```


Data Set URL: http://archive.ics.uci.edu/ml/datasets/Concrete+Compressive+Strength

***Information about the data set:***


The features in the data set are,

No| Name                    
--| ----
1 |  Cement (component 1) -- quantitative -- kg in a m3 mixture -- Input Variable 
2 |  Blast Furnace Slag (component 2) -- quantitative -- kg in a m3 mixture -- Input Variable    
3 |  Fly Ash (component 3) -- quantitative -- kg in a m3 mixture -- Input Variable 
4 |  Water (component 4) -- quantitative -- kg in a m3 mixture -- Input Variable 
5 |  Superplasticizer (component 5) -- quantitative -- kg in a m3 mixture -- Input Variable 
6 |  Coarse Aggregate (component 6) -- quantitative -- kg in a m3 mixture -- Input Variable 
7 |  Fine Aggregate (component 7)	-- quantitative -- kg in a m3 mixture -- Input Variable 
8 |  Age -- quantitative -- Day (1~365) -- Input Variable 
9 |  Concrete compressive strength -- quantitative -- MPa -- Output Variable
  

***Steps***
1. Visualize the data  
2. analyse the features  
3. correlation?  
4. PCA  
5. scaling or centring  
6. regression model  
7. test the model  
8. re do from 5

0. read the data
```{r}

cdata <- read.csv("CConcrete_Data.csv",h=T)
#check for NA
ifelse(length(which(is.na(cdata)))==0, "NO NAs available", "Yes,NAs are available")
```
  
1. Visualize the data
```{r}
pairs(cdata)
```
since this is a dataset with 1000+ records and 9 features, pairwise plot is not-so-great method to visualise. We can use a image plot of correlation matrix to find the correlation between the different features or we can hand pick the fetaures and compare them using scatter plot.  
The image plot of the correlation matrix,

***I may have to remove the y variable!!***

```{r}
C_corr <- cor(round(cdata,2))
corrplot(C_corr, method = "pie")
corrplot.mixed(C_corr, lower = "number", upper="square",order="AOE")
```

As seen in the correlation plot, (water, plasticizer) are correlated and (water, aggregate) are correlated, (cement, Fly_Ash) are correlated.

```{r}
par(mfrow=c(2,2))
attach(cdata)
plot(cdata$Water, cdata$Superplasticizer)
plot(cdata$Water,cdata$Coarse_Aggregate)
plot(cdata$Cement, cdata$Fly_Ash)
```

Next, we can do a PCA analysis to learn the features that are of relatively important compared to others.

#PCA analyis#





